---
title: "House Selling Prices"
author: "Ricardo Leon Coronado"
date: "11/16/2021"
output: html_document
---

```{r}
library(tidyverse)
library(mosaic)
library(pander)

options(scipen = 999)
```


```{r}
train <- read.csv("C:/Users/ricle/OneDrive - BYU-Idaho/Fall 21/Statistics-Notebook-master(1)/Statistics-Notebook-master/Analyses/Math425/HouseSellingPrices/train.csv", stringsAsFactors=TRUE)
View(train)
```

## The lm

```{r}
# Models

train1 <- train %>% 
  mutate(TotalSF = X1stFlrSF + X2ndFlrSF + TotalBsmtSF, 
         BathTotal = FullBath + HalfBath/2, 
         BathTotal = as.factor(BathTotal),
         BathTotal = as.factor(case_when(
           BathTotal %in% c(0, 0.5, 1, 1.5) ~ 0,
           BathTotal %in% c(2, 2.5) ~ 1,
           BathTotal %in% c(3, 3.5) ~ 2)),
         YrBuilt01 = as.factor(ifelse(YearBuilt > 1946, 1, 0)),
         FireplacesCat = as.factor(ifelse(Fireplaces > 0, 0, 1)),
         OverallQualCat = as.factor(ifelse(OverallQual %in% c(1:5), 0, 1)),
         GarageCarsCat = as.factor(ifelse(GarageCars %in% c(0:2), 0, 1)))

#table(train1$BathTotal)

#pairs(train1[c(1:10, 81)], panel = panel.smooth)
#mylm <- lm(SalePrice ~ TotalSF, data = train1) #0.6117

# # Input your lm(Y ~ ... ,data=rbdata)
# mylm <- lm(SalePrice ~ TotalSF + BathTotal, data = train1) #0.6282  #replace the "1" with your model
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF, data = train1) #0.6292 #got rid of change in intercepts and added the bathtotal with totalsf. 
# 
# #Using the YR built into the mix.
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF + YrBuilt01, data = train1) #0.6451 #Added YrBuilt into the mix as a factor. 1946><
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF + YrBuilt01 + YrBuilt01:TotalSF, data = train1) #0.6464 #Added the interaction term of YrBuilt with TotalSF into the mix.
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF + YrBuilt01:TotalSF, data = train1) #0.6467 #Dropped the YrBuilt01 alone and left the interaction term of YrBuilt with TotalSF into the mix.
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir, data = train1) #0.6536 #I added Central Air
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF, data = train1) #0.6572 #I added the Central Air:TotalSF and took away the CentralAir
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + Fireplaces, data = train1) #0.6683 #I added Fireplaces
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + Fireplaces + GarageCars:TotalSF, data = train1) #0.7348 #I added Garage Cars
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + Fireplaces + GarageCars:TotalSF + OverallQual, data = train1) #0.7918 #I added OverallQual
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + Fireplaces + GarageCars:TotalSF + OverallQual + GrLivArea, data = train1) #0.7935 #I added GrLivArea
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + Fireplaces + GarageCars:TotalSF + OverallQual + GrLivArea + GrLivArea:TotalSF, data = train1) #0.8026 #I added GrLivArea:TotalSF
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + Fireplaces + GarageCars:TotalSF + OverallQual + GrLivArea + GrLivArea:TotalSF + GarageArea, data = train1) #0.8037 #I added GarageArea
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + Fireplaces + GarageCars:TotalSF + OverallQual + GrLivArea + GrLivArea:TotalSF + GarageArea + GarageArea:TotalSF, data = train1) #0.8043 #I added GarageArea:TotalSF
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + FireplacesCat + GarageCars:TotalSF + OverallQual + GrLivArea + GrLivArea:TotalSF + GarageArea + GarageArea:TotalSF, data = train1) #0.8025 #I replaced FireplacesCat
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + Fireplaces + GarageCars:TotalSF + OverallQual + GrLivArea + GrLivArea:TotalSF + GarageArea + GarageArea:TotalSF, data = train1) #8119 #I reordered them and took out YrBuilt01
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir + CentralAir:TotalSF + Fireplaces + GarageCars:TotalSF + OverallQual + GrLivArea + GrLivArea:TotalSF + GarageArea + GarageArea:TotalSF, data = train1) #8135 #Added Central Air individually
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + FireplacesCat + GarageCars:TotalSF + OverallQual + GrLivArea + GrLivArea:TotalSF + GarageArea + GarageArea:TotalSF, data = train1) #8119 #I reordered them and took out YrBuilt01
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + FireplacesCat + GarageCars:TotalSF + OverallQualCat + GrLivArea + GrLivArea:TotalSF + GarageArea + GarageArea:TotalSF, data = train1) #7693 #I replaced OverallQualCat
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + FireplacesCat + GarageCars:TotalSF + OverallQualCat + OverallQualCat:TotalSF + GrLivArea + GrLivArea:TotalSF + GarageArea + GarageArea:TotalSF, data = train1) #0.7857 #I added OverallQualCat:TotalSF
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + FireplacesCat + GarageCars:TotalSF + OverallQualCat + OverallQualCat:TotalSF + GrLivArea + GrLivArea:TotalSF + GarageArea + GarageArea:TotalSF, data = train1) #0.7857 After trying all options I decided to stick with these.
# 
# #Lizs try
# mylm <- lm(SalePrice ~ TotalSF + BathTotal + BathTotal:TotalSF + YrBuilt01:TotalSF + CentralAir:TotalSF + FireplacesCat + GarageCars:TotalSF + OverallQualCat + OverallQualCat:TotalSF + GrLivArea + GrLivArea:TotalSF +FireplacesCat:TotalSF, data = train1) #0.7857 After trying all options I decided to stick with these.
# 
# #THIS WORKS!!!
# mylm <- lm(SalePrice ~ TotalSF + BathTotal  + GarageCars:TotalSF + OverallQualCat + OverallQualCat:TotalSF +FireplacesCat:TotalSF, data = train1) #0.7457
# 
# #mylm <- lm(SalePrice ~ TotalSF + BathTotal + BathTotal:TotalSF + GarageCars + GarageCars:TotalSF + OverallQualCat + OverallQualCat:TotalSF + FireplacesCat:TotalSF, data = train1) #0.7457
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal  + GarageCarsCat:TotalSF + OverallQualCat + OverallQualCat:TotalSF +FireplacesCat:TotalSF, data = train1) #Replaced GarageCarsCat #Not significant :(
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal  + GarageCarsCat + OverallQualCat + OverallQualCat:TotalSF +FireplacesCat:TotalSF, data = train1) #Replaced GarageCarsCat:TotalSF for only GarageCarsCat #All significant #0.7383
# 
# mylm <- lm(SalePrice ~ TotalSF + BathTotal  + GarageCarsCat + GarageCarsCat:BathTotal + OverallQualCat + OverallQualCat:TotalSF + FireplacesCat:TotalSF, data = train1) #Added GarageCarsCat:BathTotal #All significant #0.7407
 
#summary(train1$GarageCars)
# summary(mylm)
# 
# pander(summary(mylm)$coefficients)
```


```{r}
#This was the model of my choice:

mylm <- lm(SalePrice ~ TotalSF + BathTotal + GarageCarsCat + OverallQualCat + OverallQualCat:TotalSF + FireplacesCat:TotalSF, data = train1) #Added GarageCarsCat:BathTotal #All significant #0.7407

pander(summary(mylm))
```


```{r}
#Plotting graph 1

plot(SalePrice ~ TotalSF, data = train1, main = "House Selling Prices when Bathroom Total is 0 - 1.5")

b <- coef(mylm)

#Thus...
BathTotal = 0 #Small baths

GarageCarsCat = 0
OverallQualCat = 0
FireplacesCat = 0

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[1], add = TRUE)

GarageCarsCat = 1
OverallQualCat = 0
FireplacesCat = 0

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[2], add = TRUE)

GarageCarsCat = 0
OverallQualCat = 1
FireplacesCat = 0

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[3], add = TRUE)

GarageCarsCat = 0
OverallQualCat = 0
FireplacesCat = 1

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[4], add = TRUE)

GarageCarsCat = 1
OverallQualCat = 1
FireplacesCat = 0

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[5], add = TRUE)

GarageCarsCat = 0
OverallQualCat = 1
FireplacesCat = 1

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[6], add = TRUE)

GarageCarsCat = 1
OverallQualCat = 0
FireplacesCat = 1

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[7], add = TRUE)

GarageCarsCat = 1
OverallQualCat = 1
FireplacesCat = 1

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[8], add = TRUE)

legend("topleft", title = "Garage, OvQual, Fire", legend= c("(000)", "(100)", "(010)", "(001)", "(110)", "(011)", "(101)", "(111)"), bty="n", lty=1, col=c(palette()[1], palette()[2], palette()[3], palette()[4], palette()[5], palette()[6], palette()[7], palette()[8]), cex=0.8)

```


```{r}
#Plotting graph 2

plot(SalePrice ~ TotalSF, data = train1, main = "House Selling Prices when Bathroom Total is 2 - 2.5")

b <- coef(mylm)

BathTotal = 1 #Medium baths

GarageCarsCat = 0
OverallQualCat = 0
FireplacesCat = 0

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[1], add = TRUE)

GarageCarsCat = 1
OverallQualCat = 0
FireplacesCat = 0

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[2], add = TRUE)

GarageCarsCat = 0
OverallQualCat = 1
FireplacesCat = 0

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[3], add = TRUE)

GarageCarsCat = 0
OverallQualCat = 0
FireplacesCat = 1

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[4], add = TRUE)

GarageCarsCat = 1
OverallQualCat = 1
FireplacesCat = 0

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[5], add = TRUE)

GarageCarsCat = 0
OverallQualCat = 1
FireplacesCat = 1

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[6], add = TRUE)

GarageCarsCat = 1
OverallQualCat = 0
FireplacesCat = 1

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[7], add = TRUE)

GarageCarsCat = 1
OverallQualCat = 1
FireplacesCat = 1

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[8], add = TRUE)

legend("topleft", title = "Garage, OvQual, Fire", legend= c("(000)", "(100)", "(010)", "(001)", "(110)", "(011)", "(101)", "(111)"), bty="n", lty=1, col=c(palette()[1], palette()[2], palette()[3], palette()[4], palette()[5], palette()[6], palette()[7], palette()[8]), cex=0.8)

```


```{r}
#Plotting graph 3

plot(SalePrice ~ TotalSF, data = train1, main = "House Selling Prices when Bathroom Total is 3 - 3.5" )

b <- coef(mylm)

BathTotal = 2 #Large baths

GarageCarsCat = 0
OverallQualCat = 0
FireplacesCat = 0

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[1], add = TRUE)

GarageCarsCat = 1
OverallQualCat = 0
FireplacesCat = 0

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[2], add = TRUE)

GarageCarsCat = 0
OverallQualCat = 1
FireplacesCat = 0

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[3], add = TRUE)

GarageCarsCat = 0
OverallQualCat = 0
FireplacesCat = 1

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[4], add = TRUE)

GarageCarsCat = 1
OverallQualCat = 1
FireplacesCat = 0

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[5], add = TRUE)

GarageCarsCat = 0
OverallQualCat = 1
FireplacesCat = 1

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[6], add = TRUE)

GarageCarsCat = 1
OverallQualCat = 0
FireplacesCat = 1

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[7], add = TRUE)

GarageCarsCat = 1
OverallQualCat = 1
FireplacesCat = 1

curve(b[1] + b[2]*TotalSF + b[3]*BathTotal + b[4]*BathTotal + b[5]*GarageCarsCat + b[6]*OverallQualCat + b[7]*TotalSF*OverallQualCat + b[8]*TotalSF*FireplacesCat, xname = "TotalSF", col = palette()[8], add = TRUE)


legend("topleft", title = "Garage, OvQual, Fire", legend= c("(000)", "(100)", "(010)", "(001)", "(110)", "(011)", "(101)", "(111)"), bty="n", lty=1, col=c(palette()[1], palette()[2], palette()[3], palette()[4], palette()[5], palette()[6], palette()[7], palette()[8]), cex=0.8)

```


```{r}
#Validation

set.seed(121)
n <- nrow(train1)
nkeep <- sample(1:n, 1168)

my_train <- train1[nkeep, ] #1168 values goes in here (80%)
my_test <- train1[-nkeep, ] #n values go in here (the remaining 292 not keeps - 20%)
```

Interpretation:

It's interesting to notice that the starter value of a home

The higher the number of square foots the bigger the price of a home, according to our summary every additional square foot increases the price of a home by 28.35 usd. It's interesting to notice that if you add a fireplace to a home that affects the price per additional square foot which decreases by 6.754 usd. It's weird to notice as well that when the quality of a home is high the price seems to drop according to the model which could be an error somewhere in the code.
